\chapter{DYNAMICS}

We are now concerned with the dynamics of charges and currents in the presence of electromagnetic fields. We will first discuss the motion of a single charge in an electromagnetic field, and then move on to the dynamics of continuous charge and current distributions. Recall that one had the \eax{magnetic field} $\bv{B}$ which led to the \eax{magnetic force} $\bv{F}_{\text{mag}} = q \bv{v} \times \bv{B}$ on a charge $q$ moving with velocity $\bv{v}$, and the electric field $\bv{E}$ which led to the electric force $\bv{F}_{\text{elec}} = q \bv{E}$ on a charge $q$. The total force on a charge $q$ in the presence of both electric and magnetic fields is given by the \eax{Lorentz force}:
\begin{align}
\bv{F} = q (\bv{E} + \bv{v} \times \bv{B}).
\end{align}
If we look at the pure magnetic force, note that the work done is simply $\di{W} = \bv{F}_{\text{mag}} \cdot \di{\bv{l}} = \bv{F}_{\text{mag}} \cdot \bv{v} \di{t} = 0$ since $\bv{F}_{\text{mag}}$ and $\bv{v}$ are perpendicular; we can state this as a theorem.

\begin{theorem}
    Magnetic fields do no work.
\end{theorem}

Recall that we also define the \eax{electric current} $I$ as the rate of flow of charge $\di{q}/\di{t}$. We can now modify the Lorentz force law; suppose we have a wire with charge $\lambda$ per unit length flowing with velocity $\bv{v}$, then the current is given by $I = \di{q}/\di{t} = \lambda \di{l}/\di{t} = \lambda v$. From here, current is promoted to a vector $\bv{I} = \lambda \bv{v}$. Thus,
\begin{align}
    \bv{F} = \int \di{q} (\bv{v} \times \bv{B}) = \int \lambda \di{l} (\bv{v} \times \bv{B}) = \int \di{l} (\bv{I} \times \bv{B}).
\end{align}
Suppose we have a rectangular loop of wire of side length $l$ carrying a current $I$, half-immersed in a perpendicular magnetic field $\bv{B}$; it also upholds a weight $mg$ attached to the other side of the loop. The magnetic force on the loop is given by $\bv{F} = \int \di{l} (\bv{I} \times \bv{B}) = I l B$, and the weight is given by $mg$. For the loop to be in equilibrium, we must have $I l B = mg$, or $I = mg/Bl$. What happens if we increase the current $I$, leading to $I > mg/Bl?$ The loop will be lifted up, and the weight will be lifted up as well. Say it has moved upwards by a distance $h$. But doesn't this imply that $W_{\text{mag}} = IBlh$, a non-zero work done by the magnetic field? When the loop is rising, the charges in the wire also acquire a vertical component of velocity, as well as the horizontal. The magnetic force on the charges is given by $\bv{F} = q (\bv{v} \times \bv{B})$, and the work done is given by $\di{W} = \bv{F} \cdot \di{\bv{l}} = q (\bv{v} \times \bv{B}) \cdot \di{\bv{l}}$. Since $\bv{v}$ and $\di{\bv{l}}$ are parallel, we have $\di{W} = 0$. The question is, where does the work come from? In fact, to maintain the current $I$ in the loop, we need to have a power source, such as a battery, which does work on the charges to maintain the current. The work done by the battery is given by $W_{\text{battery}} = IBlh$. The mechanical analogy is of a block sliding down a frictionless incline; the normal reaction force does no work, but adds an opposing horizontal force to the block, which is balanced by the component of the weight of the block along the incline.
\\ \\
\textit{March 16th.}

Now suppose we have a surface carrying a \eax{surface current density} defined as $\bv{K} = \di{\bv{I}}/\di{l}_{\perp}$, where $\di{l}_{\perp}$ is the length element perpendicular to the direction of the current. We may also write $\bv{K} = \sigma \bv{v}$, where $\sigma$ is the charge per unit area. Hence, the force on a surface current density is given by
\begin{align}
    \bv{F} = \int (\bv{v} \times \bv{B}) \sigma \, \di{a} = \int (\bv{K} \times \bv{B})\, \di{a}.
\end{align}
Finally, we can also define a \eax{volume current density} $\bv{J} = \di{\bv{I}}/\di{a}_{\perp}$, where $\di{a}_{\perp}$ is the area element perpendicular to the direction of the current. We can also write $\bv{J} = \rho \bv{v}$, where $\rho$ is the charge per unit volume. Hence, the force on a volume current density is given by
\begin{align}
    \bv{F} = \int (\bv{v} \times \bv{B}) \rho \, \di{V} = \int (\bv{J} \times \bv{B})\, \di{V}.
\end{align}
Suppose we have a current crossing a given surface $S$. The current crossing the surface is given by $I = \int_{S} \bv{J} \cdot \di{\bv{a}} = \int_{S} J \di{a}_{\perp}$, where $\di{a}_{\perp}$ is the area element perpendicular to the direction of the current. Now suppose that the surface $S$ were closed. Then, divergence theorem tells us
\begin{align}
    \oint_{S} \bv{J} \cdot \di{\bv{a}} = \int_{\curs{v}} \nabla \cdot \bv{J} \, \di{V}.
\end{align}
Note that the current here is essentially the rate of flow of charge leaving the volume $\curs{v}$ enclosed by the surface $S$. Hence, we must have
\begin{align}
    -\frac{\di{}}{\di{t}} \int_{\curs{v}} \rho \, \di{V} = \oint_{S} \bv{J} \cdot \di{\bv{a}} = \int_{\curs{v}} \nabla \cdot \bv{J} \, \di{V}.
\end{align}
Since this must hold for any arbitrary volume $\curs{v}$, and using the fact that $\frac{\di{}}{\di{t}} \int_{\curs{v}} \rho \, \di{V} = \int_{\curs{v}} \frac{\partial{\rho}}{\partial{t}} \, \di{V}$, we get
\begin{align}
    \frac{\partial{\rho}}{\partial{t}} + \nabla \cdot \bv{J} = 0.
\end{align}
This is the \eax{continuity equation} for electric charge. It is a statement of the conservation of electric charge, and it tells us that the rate of change of charge density in a given volume is equal to the negative of the divergence of the current density. In other words, if there is a net outflow of current from a given volume, then the charge density in that volume must decrease, and vice versa. The continuity equation is a fundamental equation in electrodynamics, and it is satisfied by all physical processes involving electric charge and current.

\section{Analogous Ideas}

We wish to translate some of the laws regarding electrostatics to the case of electrodynamics, or magnetostatics. Note that just as electrostatics arises from the study of stationary charges, magnetostatics arises from the study of steady currents, since this implies $\frac{\partial{\rho}}{\partial{t}} = 0$ and $\nabla \cdot \bv{J} = 0$. We will now derive the analogue of Coulomb's law for magnetostatics, which is known as the \eax{Biot-Savart law}. It states this:
\begin{align}
    \bv{B}(\bv{r}) = \frac{\mu_{0}}{4 \pi} \int \frac{I \, \di{\bv{l}'} \times \hat{\bcurs{r}}}{\curs{r}^{2}},
\end{align}
where $\hat{\bcurs{r}}$ is the unit vector pointing from the current element $\di{\bv{l}'}$ to the point $\bv{r}$ where we want to calculate the magnetic field, and $\curs{r}$ is the distance between the current element and the point $\bv{r}$. It is the magnetic cfield produced at a point $\bv{r}$ by a current $I$ flowing through a wire. Here, $\mu_{0}$ is the \eax{permeability of free space}, with value $\mu_{0} = 4 \pi \times 10^{-7} \, \text{N/A}^2$. The units of $\bv{B}$ are thus Tesla (T), where $1 \, \text{T} = 1 \, \text{N/A m}$. It can be seen as an inverse-square law phenomenon, but is actually more complex than that, since the direction of the magnetic field is given by the cross product $\di{\bv{l}'} \times \hat{\bcurs{r}}$.
\\

As a concrete example, let us calculate the magnetic field at a distance of $s$ from a long straight wire carrying a current $I$. We can choose our coordinate system such that the wire lies along the $z$-axis, and the point where we want to calculate the magnetic field is at a distance $s$ from the wire in the $x$-direction. The current element $\di{\bv{l}'}$ can be written as $\di{z'} \hat{\bv{z}}$, and the unit vector $\hat{\bcurs{r}}$ can be written as $\frac{s \hat{\bv{x}} - z' \hat{\bv{z}}}{\sqrt{s^2 + z'^2}}$. Hence, we have
\begin{align}
\bv{B}(\bv{r}) &= \frac{\mu_{0}}{4 \pi} \int_{-\infty}^{\infty} \frac{I \, \di{z'} \hat{\bv{z}} \times \frac{s \hat{\bv{x}} - z' \hat{\bv{z}}}{\sqrt{s^2 + z'^2}}}{s^2 + z'^2} = \frac{\mu_{0} I s}{4 \pi} \int_{-\infty}^{\infty} \frac{\di{z'}}{(s^2 + z'^2)^{3/2}} \hat{\bv{y}} = \frac{\mu_{0} I}{2 \pi s} \hat{\bv{y}}.
\end{align}
From here, we can calculate the force between two long straight wires carrying currents $I_{1}$ and $I_{2}$, separated by a distance $d$. The magnetic field produced by the first wire at the location of the second wire is given by $B = \frac{\mu_{0} I_{1}}{2 \pi d}$. The force on the second wire due to this magnetic field is given by $F = I_{2}B \int \di{l}$. The second wire is also infinitely long, so we can instead write the force per unit length as
\begin{align}
    f = \frac{\di{F}}{\di{l}} = I_{2}B = \frac{\mu_{0} I_{1} I_{2}}{2 \pi d}.
\end{align}
Let us find the divergence of the magnetic field. A little bit of vector calculus gives us
\begin{align}
    \nabla \cdot \bv{B}(\bv{r}) = \frac{\mu_{0}}{4 \pi} \int \nabla \cdot \left( \frac{\bv{J} \times \hat{\bcurs{r}}}{\curs{r}^{2}} \right) \, \di{V} = 0,
\end{align}
since $\nabla \cdot (\bv{A} \times \bv{B}) = \bv{B} \cdot (\nabla \times \bv{A}) - \bv{A} \cdot (\nabla \times \bv{B})$, and $\nabla \times \hat{\bcurs{r}} = 0$ and $\nabla \times \bv{J} = 0$ for steady currents. Hence, we have the following result: The magnetic field is divergence-free, or equivalently, there are no magnetic monopoles. This is a fundamental result in magnetostatics, and it tells us that the magnetic field lines are always closed loops, and they never begin or end at any point in space. This is in contrast to the electric field, which can have sources and sinks (positive and negative charges). The absence of magnetic monopoles is a fundamental symmetry of nature, and it has been confirmed by numerous experiments. For the curl of the magnetic field, vector calculus again gives us
\begin{align}
    \nabla \times \bv{B}(\bv{r}) = \frac{\mu_{0}}{4 \pi} \int \nabla \times \left( \frac{\bv{J} \times \hat{\bcurs{r}}}{\curs{r}^{2}} \right) \, \di{V} = \mu_{0} \bv{J}(\bv{r}),
\end{align}
where we have used the fact that $\nabla \times \left( \frac{\hat{\bcurs{r}}}{\curs{r}^{2}} \right) = 4 \pi \delta(\bv{r} - \bv{r}')$. This is known as \eax{Ampère's law}, and it tells us that the curl of the magnetic field is proportional to the current density. In other words, currents are the sources of magnetic fields, just as charges are the sources of electric fields. If we integrate both sides of Ampère's law over a surface $S$ with boundary $\partial S$, we get
\begin{align}
    \oint \bv{B} \cdot \di{\bv{l}} = \int_{S} (\nabla \times \bv{B}) \cdot \di{\bv{a}} = \mu_{0} \int_{S} \bv{J} \cdot \di{\bv{a}} = \mu_{0} I_{\text{enc}},
\end{align}
where $I_{\text{enc}}$ is the current enclosed by the loop $\partial S$. This is the integral form of Ampère's law, and it is often more useful for calculating magnetic fields in situations with high symmetry.
\\ \\
\noindent \textit{March 18th.}

We had already derived the current due to a long straight wire, which is given by $B = \frac{\mu_{0} I}{2 \pi s}$. The problem becomes almost trivial using Ampère's law. We can choose an Amperian loop to be a circle of radius $s$ centered on the wire. The magnetic field is tangent to the loop at every point, and its magnitude is constant along the loop. Thus, we have
\begin{align}
\oint \bv{B} \cdot \di{\bv{l}} = B \oint \di{l} = B (2 \pi s) = \mu_{0} I \implies B = \frac{\mu_{0} I}{2 \pi s}.
\end{align}
Now consider a \eax{solenoid}, which is a long cylindrical coil of wire with $n$ turns per unit length, carrying a current $I$. With a little bit of thought, we can see that the magnetic field inside the solenoid is uniform and parallel to the axis of the solenoid. We can choose an Amperian loop to be a rectangle that extends inside and outside the solenoid, with one side parallel to the axis of the solenoid and the other side perpendicular to it. The magnetic field outside the solenoid is negligible, so the contribution to the line integral from the outside part of the loop is zero. The contribution from the inside part of the loop is $B \cdot l$, where $l$ is the length of the side of the loop that is parallel to the axis of the solenoid. The current enclosed by the loop is $I_{\text{enc}} = n I l$, since there are $n$ turns per unit length and each turn carries a current $I$. Hence, we have
\begin{align}
B \cdot l = \mu_{0} n I l \implies B = \mu_{0} n I.
\end{align}

\subsection{Potential}

We have already seen that $\nabla \cdot \bv{E} = \frac{\rho}{\epsilon_{0}}$ and $\nabla \times \bv{E} = 0$ in electrostatics, and $\nabla \cdot \bv{B} = 0$ and $\nabla \times \bv{B} = \mu_{0} \bv{J}$ in magnetostatics. We can now ask if there are potentials associated with the electric and magnetic fields. In electrostatics, since $\nabla \times \bv{E} = 0$, we can write $\bv{E} = -\nabla V$, where $V$ is the electric potential. In magnetostatics, since $\nabla \cdot \bv{B} = 0$, we can write $\bv{B} = \nabla \times \bv{A}$, where $\bv{A}$ is the \eax{magnetic vector potential}. Note that just like $V$ was arbitrary up to an additive constant, $\bv{A}$ is also arbitrary up to the gradient of a scalar function, since $\nabla \times (\bv{A} + \nabla \lambda) = \nabla \times \bv{A}$.
\\ \\
\textit{March 23rd.}

We can exploit the freedom in the definition of $\bv{A}$ to choose a particular `gauge'. One common choice is to impose the condition $\nabla \cdot \bv{A} = 0$. To do so, suppose we have a $\bv{A}_{0}$ which is not divergence free. We can write $\bv{A}_{0} = \bv{A} + \nabla \lambda$, where $\nabla \cdot \bv{A} = 0$. Hence, we have $\nabla \cdot \bv{A}_{0} = \nabla \cdot (\bv{A} + \nabla \lambda) = \nabla^{2} \lambda$. Thus, we can solve the Poisson equation $\nabla^{2} \lambda = \nabla \cdot \bv{A}_{0}$ to find $\lambda$, and hence find a divergence-free vector potential $\bv{A}$. Thus, we have
\begin{align}
    \lambda = \frac{1}{4 \pi} \int \frac{\nabla \cdot \bv{A}_{0}(\bv{r}')}{\curs{r}} \, \di{V}'.
\end{align}
If $\nabla \cdot \bv{A}_{0}$ does not go to zero at infinity, we have to find other means of solving the equation. Taking the curl of both sides of $\bv{B} = \nabla \times \bv{A}$, we get $\nabla \times \bv{B} = \nabla \times (\nabla \times \bv{A})$. Using the vector identity $\nabla \times (\nabla \times \bv{A}) = \nabla (\nabla \cdot \bv{A}) - \nabla^{2} \bv{A}$, and using the gauge condition $\nabla \cdot \bv{A} = 0$, we get $\nabla \times \bv{B} = -\nabla^{2} \bv{A}$. Using Ampère's law, we have $\mu_{0} \bv{J} = -\nabla^{2} \bv{A}$, which is a Poisson equation for each component of $\bv{A}$. Thus, this gives
\begin{align}
    \bv{A} = \frac{\mu_{0}}{4 \pi} \int \frac{\bv{J}(\bv{r}')}{\curs{r}} \, \di{V}'.
\end{align}
For line currents and surface currents, we can write $\bv{A}$ as
\begin{align}
    \bv{A} = \frac{\mu_{0}}{4 \pi} \int \frac{\bv{I} \, \di{\bv{l}'}}{\curs{r}} \quad \text{and} \quad \bv{A} = \frac{\mu_{0}}{4 \pi} \int \frac{\bv{K}(\bv{r}')}{\curs{r}} \, \di{a}',
\end{align}
respectively. 

\subsubsection{Magnetostatic Boundary Conditions}

Recall that $\bv{E}$ was discontinuous at surface charges. Analogously, $\bv{B}$ is discontinuous at surface currents. Using the fact that $\nabla \cdot \bv{B} = 0$ and $\oint \bv{B} \cdot \di{\bv{a}} = 0$, we have $\oint \bv{B} \cdot \di{\bv{l}} = \mu_{0} I_{\text{enc}}$, where $I_{\text{enc}}$ is the current enclosed by the loop. So, $B^{\parallel}_{\text{above}} - B^{\parallel}_{\text{below}} = \mu_{0} K$, where $K$ is the surface current density. Combining gives
\begin{align}
    \bv{B}_{\text{above}} - \bv{B}_{\text{below}} = \mu_{0} \bv{K} \times \hat{\bv{n}},
\end{align}
where $\hat{\bv{n}}$ is the unit normal vector pointing from the below region to the above region. And just like before, $\bv{A}_{\text{above}} - \bv{A}_{\text{below}} = 0$ since $\nabla \cdot \bv{A} = 0$ and $\oint \bv{A} \cdot \di{\bv{l}} = 0$. We can also attempt a multipole expansion of the vector potential $\bv{A}$, just like we did for the scalar potential $V$. We have
\begin{align}
    \frac{1}{\curs{r}} = \frac{1}{\sqrt{r^{2}+r'^{2}+2rr'\cos \theta}} = \frac{1}{r} \sum_{n = 0}^{\infty} \left( \frac{r'}{r} \right)^{n} P_{n}(\cos \theta),
\end{align}
so
\begin{align}
    \bv{A} = \frac{\mu_{0} I}{4 \pi} \oint \frac{\di{\bv{l}'}}{\curs{r}} = \frac{\mu_{0} I}{4 \pi r} \sum_{n = 0}^{\infty} \frac{1}{r^{n}} \oint r'^{n} P_{n}(\cos \theta) \, \di{\bv{l}'}.
\end{align}
The vector potential of an infinite solenoid carryinng current $I$ and with $n$ turns per unit length, with radius $R$, is given by
\begin{align}
    \bv{A} = \frac{\mu_{0}}{4 \pi} \int \frac{\bv{J}(\bv{r}')}{\curs{r}} \, \di{V}'.
\end{align}

\section{The Electromotive Force}
\textit{March 25th.}

Plainly stated, \eax{Ohm's law} states that the current density $\bv{J}$ is proportional to the force per unit charge $\bv{f}$, with the constant of proportionality being the \eax{conductivity} $\sigma$ of the material. Hence, we have
\begin{align}
    \bv{J} \propto \bv{f} \implies \bv{J} = \sigma \bv{f}.
\end{align}
In the presence of an electric field $\bv{E}$ and a magnetic field $\bv{B}$, the force per unit charge is given by $\bv{f} = \bv{E} + \bv{v} \times \bv{B}$. In practice, however, the magnetic force is negligible compared to the electric force, since the velocity of the charges is typically very small. Hence, we can write Ohm's law as
\begin{align}
    \bv{J} = \sigma \bv{E}.
\end{align}
This is not the form we are familiar with; to that end, let us take the divergence of both sides. For materials with constant conductivity, we have $\nabla \cdot \bv{J} = \sigma \nabla \cdot \bv{E}$. Using Gauss's law, we have $\nabla \cdot \bv{E} = \frac{\rho}{\epsilon_{0}}$, so $\nabla \cdot \bv{J} = \frac{\sigma}{\epsilon_{0}} \rho$. Therefore, for study currents, $\nabla \cdot \bv{J} = 0$, so $\rho = 0$. That is, there is no net charge density within the volume of the conductor and all charge must reside on the surface.

In a typical circuit, a battery maintains an \eax{electromotive force} (emf), which is the work done by the battery per unit charge to move the charges through the circuit. The force per unit charge in this case is $\bv{f} = \bv{f}_{s} + \bv{E}$, where $\bv{f}_{s}$ is the force per unit charge due to the battery. The net effect is summarized by following quantity,
\begin{align}
    \cE = \oint \bv{f} \cdot \di{\bv{l}} = \oint \bv{f}_{s} \cdot \di{\bv{l}} + \oint \bv{E} \cdot \di{\bv{l}} = \oint \bv{f}_{s} \cdot \di{\bv{l}},
\end{align}
where we have used the fact that $\oint \bv{E} \cdot \di{\bv{l}} = 0$ for electrostatic fields. The quantity $\cE$ is the emf of the circuit. For an ideal source, $\bv{f} = 0$ or $\bv{f}_{s} = -\bv{E}$. Recall that $V = -\int \bv{E} \cdot \di{\bv{l}}$, so $\cE = V$. Thus, it is valid to intepret $\cE$ as the potential difference across the battery too.

Now suppose we have a rectangular loop of wire of height $h$ being pulled out of a uniform and perpendicular magnetic field $\bv{B}$ with a velocity $\bv{v}$. Note that $\bv{f}_{\text{pull}}$, the force due to the pulling, contributes nothing to the emf, since $\oint \bv{f}_{\text{pull}} \cdot \di{\bv{l}} = 0$. The only contribution to the emf is from the magnetic force, which is given by $\oint \bv{f}_{\text{mag}} \cdot \di{\bv{l}} = \oint (\bv{v} \times \bv{B}) \cdot \di{\bv{l}} = vBh$. Hence, we have $\cE = vBh$. This is known as the \eax{motional emf}, and it is a consequence of the fact that the magnetic force on the charges in the wire does work as the wire moves through the magnetic field. An alternative way of computing the motional emf is to note that the \eax{magnetic flux} through the loop is given by $\Phi = \int \bv{B} \cdot \di{\bv{a}} = Bhx$, where $x$ is the length of the loop that is still within the magnetic field. Hence, we have $\cE = -\frac{\di{\Phi}}{\di{t}} = vBh$, which is consistent with our previous calculation. Note that this result of
\begin{align}
    \cE = -\frac{\di{\Phi}}{\di{t}}
\end{align}
generalizes to loops of arbitrary shapes, non-constant magnetic fields, and non-uniform velocities. There is some ambiguity in the fact that in $\int \bv{B} \cdot \di{\bv{a}}$, which direction do we choose for the area element $\di{\bv{a}}$? The standard choice is to use the right-hand rule: if the current in the loop is flowing in a counterclockwise direction, then the area element points upwards; if the current is flowing in a clockwise direction, then the area element points downwards. The equation $\cE = -\frac{\di{\Phi}}{\di{t}}$ is known as \eax{Faraday's law of induction}, and it tells us that a changing magnetic flux through a loop induces an emf in the loop. Note that
\begin{align}
    \cE = \oint \bv{E} \cdot \di{\bv{l}} = -\frac{\di{\Phi}}{\di{t}} = -\int \frac{\partial{\bv{B}}}{\partial{t}} \cdot \di{\bv{a}},
\end{align}
so $\oint \bv{E} \cdot \di{\bv{l}} = -\int \frac{\partial{\bv{B}}}{\partial{t}} \cdot \di{\bv{a}}$. Using Stokes' theorem, we have $\int (\nabla \times \bv{E}) \cdot \di{\bv{a}} = -\int \frac{\partial{\bv{B}}}{\partial{t}} \cdot \di{\bv{a}}$, so
\begin{align}
    \nabla \times \bv{E} = -\frac{\partial{\bv{B}}}{\partial{t}}.
\end{align}
This is the differential form of Faraday's law.

\subsection{Induction}

\noindent \textit{March 30th.}

Faraday's law may be used in many different ways. For example, suppose a uniform magnetic field $\bv{B}$ is passing perpendicularly. If the magnetic field is changing with time, then there will be an electric field induced. For an amperian loop that is a circle of radius $s$, we have
\begin{align}
    E \cdot 2 \pi s = - \frac{\di{\Phi}}{\di{t}} = -\pi s^{2} \frac{\di{B}}{\di{t}} \implies E = -\frac{s}{2} \frac{\di{B}}{\di{t}} \implies \bv{E} = -\frac{s}{2} \frac{\di{B}}{\di{t}} \hat{\bvs{\phi}},
\end{align}

Suppose we have a freely, and horizontally, suspended circular disc of inner radius $a$ and outer radius $b$, with a linear charge $\lambda$ per unit length glued to the outer circumference. Suppose also that a uniform magnetic field $\bv{B}_{0}$ passes through the middle perpendicularly upwards. We now switch off $B_{0}$ instantaneously. What happens? Switching off $B_{0}$ induces an electric field $\bv{E}$ along the $\hat{\bvs{\phi}}$ direction. But there is also the line charge $\lambda$ along the rim; a current is induced which moves the charges and the disc starts to rotate. We have
\begin{align}
    \cE = \oint \bv{E} \cdot \di{\bv{l}} = -\frac{\di{B}}{\di{t}} \pi a^{2}.
\end{align}
The torque is given by $\bv{N} = \bv{r} \times \bv{F}$, or $\di{\bv{N}} = \bv{r} \times \di{\bv{F}}$. Using $\di{F} = E \lambda \, \di{l}$, we have $\di{N} = E \lambda \, \di{l} \, b$. Integrating gives us
\begin{align}
    N = -\lambda b \pi a^{2} \frac{\di{B}}{\di{t}}.
\end{align}
The angular momentum is, thus,
\begin{align}
    N = \frac{\di{L}}{\di{t}} = - \pi a^{2} \lambda b \frac{\di{B}}{\di{t}} \implies L = -\pi a^{2} \lambda b \int_{B_{0}}^{0} \di{B} = \pi a^{2} \lambda b B_{0}.
\end{align}

Now suppose we have two loops of wire, where one carries a current $I_{1}$. The magnetic field produced is
\begin{align}
    \bv{B}_{1} = \frac{\mu_{0}I_{1}}{4 \pi} \oint \frac{\di{\bv{l}_{1}} \times \hat{\bcurs{r}}}{\curs{r}^{2}}.
\end{align}
The important thing to note is the fact that $\abs{\bv{B}_{1}} \propto I_{1}$. The magnetic flux through the second loop is given by $\Phi_{21} = \int \bv{B}_{1} \cdot \di{\bv{a}_{2}}$, so $\Phi_{21} \propto I_{1}$. This gives us $\Phi_{21} = M_{21} I_{1}$, where $M_{21}$ is the \eax{mutual inductance} between the two loops. Using the fact that $\bv{B}_{1} = \nabla \times \bv{A}_{1}$, we have $\Phi_{21} = \int (\nabla \times \bv{A}_{1}) \cdot \di{\bv{a}_{2}}$. Using Stokes' theorem, we have $\Phi_{21} = \oint \bv{A}_{1} \cdot \di{\bv{l}_{2}}$. Substituting the expression for $\bv{A}_{1}$ gives us
\begin{align}
    \Phi_{21} = \frac{\mu_{0} I_{1}}{4 \pi} \oint \oint \frac{\di{\bv{l}_{1}} \cdot (\di{\bv{l}_{2}} \times \hat{\bcurs{r}})}{\curs{r}^{2}} \implies M_{21} = \frac{\mu_{0}}{4 \pi} \oint \oint \frac{\di{\bv{l}_{1}} \cdot (\di{\bv{l}_{2}})}{\curs{r}}.
\end{align}
We observe that $M_{21}$ depends only on the geometry of the two loops, and not on the current $I_{1}$. By symmetry, we also conclude $M_{21} = M_{12} = M$.

Now suppose we have two concentric solenoids, with the inner solenoid having $n_{1}$ turns per unit length and radius $a$ and carrying a current $I$, and the outer solenoid having $n_{2}$ turns per unit length and radius $b$. Also, the inner solenoid is shorter than the outer one. We are to find the flux and the mutual inductance. Note that $B$ for the short solenoid is messy, but $B$ for the long solenoid is simple. We can exploit the symmetry of $M_{21} = M_{12}$ and change the problme; suppose the long solenoid was carrying current $I$, and the short solenoid was not carrying any current. The magneti field produced by the long solenoid is given by $\bv{B} = \mu_{0} n_{2} I \hat{\bv{z}}$. The flux through the short solenoid is given by $\Phi = \int \bv{B} \cdot \di{\bv{a}} = \pi a^{2} \mu_{0} n_{2} I$. Hence, we have $M = \frac{\Phi}{I} = \pi a^{2} \mu_{0} n_{2}$.

Suppose we have a single loop of wire with a current $I$. If we allow $I$ to vary with time, then the flux canges which induces an emf in the loop that opposes back the change in flux. This is known as \eax{self-induction}. In such a case, we have $\Phi = LI$, where $L$ is the \eax{self-inductance} of the loop. The back emf is given by $\cE = -\frac{\di{\Phi}}{\di{t}} = -L \frac{\di{I}}{\di{t}}$.

Finally, now suppose we have a circuit with a battery of emf $\cE_{0}$ and an \eax{inductor} (a device that resists changes in current) with self-inductance $L$, and a \eax{resistor} (a device that resists the flow of current) with \eax{resistance} $R$ (recall that the standard form of Ohm's law is $V = IR$, where $V$ is the potential difference across the resistor, $I$ is the current through the resistor, and $R$ is the resistance of the resistor). We are to find the current $I$ as a function of time. Adding up the emfs gives us
\begin{align}
    \cE_{0} - L \frac{\di{I}}{\di{t}} - IR = 0 \implies L \frac{\di{I}}{\di{t}} + IR = \cE_{0}.
\end{align}
This is a first-order linear ordinary differential equation, and can be easily solved using integrating factors. We have
\begin{align}
    I = \frac{\cE_{0}}{R} + ke^{-\frac{R}{L}t}
\end{align}
where $k$ is a constant that can be determined by the initial condition. If we assume that the current is zero at time $t = 0$, then we have $k = -\frac{\cE_{0}}{R}$. Hence, we have
\begin{align}
    I = \frac{\cE_{0}}{R} \left( 1 - e^{-\frac{R}{L}t} \right).
\end{align}
Here, $\tau = L/R$ is known as the \eax{time constant} of the circuit, and it characterizes how quickly the current reaches its steady-state value of $\cE_{0}/R$. For $t \gg \tau$, we have $I \approx \cE_{0}/R$, which is the steady-state current. We can think of $L$ as playing the role of `mass'. What if we also add a capacitor (a device that stores electric charge) with \eax{capacitance} $C$ (recall that the standard form of the relationship between charge and voltage for a capacitor is $Q = CV$, where $Q$ is the charge stored in the capacitor, $C$ is the capacitance of the capacitor, and $V$ is the potential difference across the capacitor)? The equation becomes
\begin{align}
    \cE_{0} = L \frac{\di{I}}{\di{t}} + IR + \frac{q}{C} = L \frac{\di{}^{2}q}{\di{t}^{2}} + R \frac{\di{q}}{\di{t}} + \frac{q}{C}.
\end{align}
Compare this with $m \frac{\di{}^{2}x}{\di{t}^{2}} + b \frac{\di{x}}{\di{t}} + kx = F_{0}$, which is the equation of motion for a damped harmonic oscillator with mass $m$, damping coefficient $b$, spring constant $k$, and external force $F_{0}$. We see that the charge $q$ on the capacitor behaves like the position $x$ of a damped harmonic oscillator, with the inductance $L$ playing the role of mass, the resistance $R$ playing the role of damping, and the inverse of capacitance $1/C$ playing the role of spring constant.

\subsubsection{Energy in Magnetic Fields}
\textit{April 1st.}

The simple idea is that building up a current takes work against the back emf $\cE = -L \frac{\di{I}}{\di{t}}$. The power is given by $P = -\cE I = L I \frac{\di{I}}{\di{t}}$. Integrating gives us the energy stored in the magnetic field of the inductor, which is given by
\begin{align}
    W = \frac{1}{2} LI^{2}.
\end{align}
The corresponding electrostatic equation was $W = \frac{1}{2} CV^{2}$. Note that $L$ depends only on the geometry. The magnetic flux in this case is
\begin{align}
    \Phi = \int_{S} \bv{B} \cdot \di{\bv{a}} = \int_{S} (\nabla \times \bv{A}) \cdot \di{\bv{a}} = \oint_{P} \bv{A} \cdot \di{\bv{l}} = LI,
\end{align}
so we can rewrite the work as
\begin{align}
    W = \frac{1}{2} \oint I\bv{A} \cdot \di{\bv{l}} = \frac{1}{2} \oint \bv{A} \cdot \bv{I} \, \di{l}.
\end{align}
This can be generalized to a volume current distribution as
\begin{align}
    W = \frac{1}{2} \int \bv{A} \cdot \bv{J} \, \di{V}.
\end{align}
Now, $\bv{J}$ can be replaced by $\frac{1}{\mu_{0}} \nabla \times \bv{B}$ using Ampère's law, so
\begin{align}
    W = \frac{1}{2 \mu_{0}} \int \bv{A} \cdot (\nabla \times \bv{B}) \, \di{V} &= \frac{1}{2 \mu_{0}} \int_{V} \bv{B} \cdot (\nabla \times \bv{A}) \, \di{V'} - \frac{1}{2 \mu_{0}} \int_{V} \nabla \cdot (\bv{A} \times \bv{B}) \, \di{V'} \notag \\ &= \frac{1}{2 \mu_{0}} \int_{V} B^{2} \, \di{V'} - \frac{1}{2 \mu_{0}} \oint_{S} (\bv{A} \times \bv{B}) \cdot \di{\bv{a}}.
\end{align}
THe second of these terms is the boundary term and is hence zero. Thus, we have
\begin{align}
    W = \frac{1}{2 \mu_{0}} \int_{V} B^{2} \, \di{V'}.
\end{align}
The corresponding electrostatic equation was $W = \frac{\epsilon_{0}}{2} \int_{V} E^{2} \, \di{V'}$. 

\section{The Laws of Electromagnetics}
Along with the Lorentz force law $\bv{f} = \bv{E} + \bv{v} \times \bv{B}$, we have the following four equations that govern all of electromagnetics:
\begin{align}
\nabla \cdot \bv{E} &= \frac{\rho}{\epsilon_{0}} \quad \text{(Gauss's law)}, \notag \\
\nabla \cdot \bv{B} &= 0 \quad \text{(no magnetic monopoles)}, \notag \\
\nabla \times \bv{E} &= -\frac{\partial{\bv{B}}}{\partial{t}} \quad \text{(Faraday's law)}, \notag \\
\nabla \times \bv{B} &= \mu_{0} \bv{J} \quad \text{(Ampère's law)}. \notag
\end{align}
Out of all of these equations, however, one has a fundamental issue. Ampère's law, in its original form, was $\nabla \times \bv{B} = \mu_{0} \bv{J}$. This is fine for steady currents, but it fails for time-varying currents. To notice this inconsistency, we can take the divergence of both sides. We have
\begin{align}
\nabla \cdot (\nabla \times \bv{B}) = \nabla \cdot (\mu_{0} \bv{J}) \implies 0 = \mu_{0} \nabla \cdot \bv{J}.
\end{align}
The last equation implies $\nabla \cdot \bv{J} = 0$, which holds only in the case of steady currents. Thus, this law must be modified to resolve this inconsistency. To do so, consider a circuit consisting of a battery of emf $\cE$ and a capacitor with capacitance $C$. The integral form of Ampère's law states that $\oint \bv{B} \cdot \di{\bv{l}} = \mu_{0} I_{\text{enc}}$, where $I_{\text{enc}}$ is the current enclosed by the loop. If we take the loop around the wire, then $I_{\text{enc}} = I$, where $I$ is the current through the wire. If we take the loop around the capacitor, then $I_{\text{enc}} = 0$, since there is no current through the capacitor. This clearly cannot be the case.

The continuity equation states that $\frac{\partial \rho}{\partial t} + \nabla \cdot \bv{J} = 0$. Using this, we have
\begin{align}
    \nabla \cdot (\nabla \times \bv{B}) = \mu_{0} \nabla \cdot \bv{J} = -\mu_{0} \frac{\partial \rho}{\partial t} = -\mu_{0} \epsilon_{0} \frac{\partial}{\partial t} (\nabla \cdot \bv{E}) = -\mu_{0} \epsilon_{0} \nabla \cdot \left( \frac{\partial{\bv{E}}}{\partial{t}} \right).
\end{align}
Hence, Faraday added a term $\mu_{0} \epsilon_{0} \frac{\partial{\bv{E}}}{\partial{t}}$ to the right-hand side of Ampère's law, so that the modified Ampère's law is given by
\begin{align}
    \nabla \times \bv{B} = \mu_{0} \bv{J} + \mu_{0} \epsilon_{0} \frac{\partial{\bv{E}}}{\partial{t}}.
\end{align}
This, along with the other three equations, are known as \eax{Maxwell's equations}, and they govern all of electromagnetics. From this modified equation, we also infer that changing electric fields can also produce magnetic fields, just as changing magnetic fields can produce electric fields. Coming back to the capacitor example, we see that
\begin{align}
    \oint \bv{B} \cdot \di{\bv{l}} = \mu_{0} \epsilon_{0} \int \frac{\partial \bv{E}}{\partial t} \cdot \di{\bv{a}} = \mu_{0} \epsilon_{0} \frac{I}{\epsilon_{0}A} \cdot A = \mu_{0} I,
\end{align}
where we used the fact that $\frac{\partial E}{\partial t} = \frac{I}{\epsilon_{0}A}$, since $I = \frac{\di{Q}}{\di{t}}$ and $Q = \epsilon_{0} EA$. Thus, the modified Ampère's law resolves the inconsistency in the original Ampère's law.
\\ \\
\textit{April 6th.}

Now suppose we have a region of space with no charges and no currents. Then, Maxwell's equations reduce to
\begin{align}
\nabla \cdot \bv{E} = 0, \quad \nabla \cdot \bv{B} = 0, \quad \nabla \times \bv{E} = -\frac{\partial{\bv{B}}}{\partial{t}}, \quad \nabla \times \bv{B} = \mu_{0} \epsilon_{0} \frac{\partial{\bv{E}}}{\partial{t}}.
\end{align}

\subsection{Poynting's Theorem}
The equations of work done,
\begin{align}
    W = \frac{\epsilon_{0}}{2} \int_{V} E^{2} \, \di{V'} \quad \text{and} \quad W = \frac{1}{2 \mu_{0}} \int_{V} B^{2} \, \di{V'},
\end{align}
suggest that the energy density of the electromagnetic field is given by
\begin{align}
    u = \frac{\epsilon_{0}}{2} E^{2} + \frac{1}{2 \mu_{0}} B^{2}.
\end{align}
Now suppose we have a charge and current configuration generating fields $\bv{E}(\bv{r},t)$ and $\bv{B}(\bv{r},t)$. The work done by the electromagnetic force is
\begin{align}
    \di{W} = \bv{F} \cdot \di{\bv{l}} = q(\bv{E} + \bv{v} \times \bv{B}) \cdot \bv{v} \, \di{t}.
\end{align}
Using $q = \rho \, \di{V}'$ and $\bv{J} = \rho \bv{v}$, we have
\begin{align}
    \frac{\di{\tilde{W}}}{\di{t}} = \int \rho (\bv{J} \cdot \bv{E}) \, \di{V}'
\end{align}
where $\di{\tilde{W}}/\di{t}$ represents the work done per unit time across a given volume $V$. We can replace $\bv{J}$ from Maxwell's equations as
\begin{align}
    \bv{E} \cdot \bv{J} = \bv{E} \cdot \left( \frac{1}{\mu_{0}} (\nabla \times \bv{B}) - \epsilon_{0} \frac{\partial \bv{E}}{\partial t} \right).
\end{align}
Using the vector identity $\nabla \cdot (\bv{E} \times \bv{B}) = \bv{B} \cdot (\nabla \times \bv{E}) - \bv{E} \cdot (\nabla \times \bv{B})$, we have
\begin{align}
    \bv{E} \cdot \bv{J} = \frac{1}{\mu_{0}} \bv{B} \cdot (\nabla \times \bv{E}) - \frac{1}{\mu_{0}} \nabla \cdot (\bv{E} \times \bv{B}) - \epsilon_{0} \bv{E} \cdot \frac{\partial \bv{E}}{\partial t}.
\end{align}
We can now replace $\nabla \times \bv{E}$ from Maxwell's equations as
\begin{align}
    \bv{E} \cdot \bv{J} &= -\frac{1}{\mu_{0}} \bv{B} \cdot \frac{\partial \bv{B}}{\partial t} - \epsilon_{0} \bv{E} \cdot \frac{\partial \bv{E}}{\partial t} - \frac{1}{\mu_{0}} \nabla \cdot (\bv{E} \times \bv{B}) \notag \\
    &= -\frac{\partial}{\partial t} \left( \frac{1}{2} \left( \epsilon_{0} E^{2} + \frac{1}{\mu_{0}} B^{2} \right) \right) - \frac{1}{\mu_{0}} \nabla \cdot (\bv{E} \times \bv{B}).
\end{align}
Therefore,
\begin{align}
    \frac{\di{\tilde{W}}}{\di{t}} = -\int_{V} \frac{\partial u}{\partial t} \, \di{V}' - \int_{V} \nabla \cdot \bv{S} \, \di{V}',
\end{align}
where $u = \frac{\epsilon_{0}}{2} E^{2} + \frac{1}{2 \mu_{0}} B^{2}$ is defined as the \eax{energy density} of the electromagnetic field, and $\bv{S} = \frac{1}{\mu_{0}} \bv{E} \times \bv{B}$ is defined as the \eax{Poynting vector}. Note that the Poynting vector $\bv{S}$ is orthogonal to both $\bv{E}$ and $\bv{B}$. Therefore, using $W$ instead of $\tilde{W}$ to represent the total work done, we have
\begin{align}
    \frac{\di{W}}{\di{t}} = -\frac{\di{U}}{\di{t}} - \oint_{S} \bv{S} \cdot \di{\bv{a}},
\end{align}
where $U$ is the total energy stored in the electromagnetic field, and $S$ is the surface bounding the volume $V$. This is known as \eax{Poynting's theorem}. Thus, the work done on charges by the electromagnetic field within a certain volume equates to a decrease in the energy stored in the electromagnetic field within that volume, plus a net outward flow of energy across the boundary of that volume. The Poynting vector $\bv{S}$ represents the energy flux density of the electromagnetic field, and its direction indicates the direction of energy flow.

Note that work done on the charges will also increase their mechanical energy. If we employ $U_{em}$ instead of $U$ to explicitly represent the energy stored in the electromagnetic field, and $u_{em}$ instead of $u$ to explicitly represent the energy density of the electromagnetic field, and finally $u_{mech}$ to represent the mechanical energy density of the charges, then we can rewrite Poynting's theorem as
\begin{align}
    \frac{\di{}}{\di{t}} \int u_{mech} \, \di{V}' = -\int \frac{\partial u_{em}}{\partial t} \, \di{V}' - \oint_{S} \bv{S} \cdot \di{\bv{a}} \implies \frac{\di{}}{\di{t}} \int (u_{mech} + u_{em}) \, \di{V}' = -\oint_{S} \bv{S} \cdot \di{\bv{a}}.
\end{align}
If we now use $u_{total} = u_{mech} + u_{em}$ to represent the total energy density, then we have
\begin{align}
    \frac{\di{}}{\di{t}} \int u_{total} \, \di{V}' = -\oint_{S} \bv{S} \cdot \di{\bv{a}} = -\int_{V} \nabla \cdot \bv{S} \, \di{V}' \implies \frac{\partial u_{total}}{\partial t} + \nabla \cdot \bv{S} = 0
\end{align}
which is the differential form of Poynting's theorem. This equation, along with
\begin{align}
    \frac{\partial \rho}{\partial t} + \nabla \cdot \bv{J} = 0
\end{align}
which is the continuity equation, are the two conservation laws of electromagnetics. The first one states that the total energy (mechanical plus electromagnetic) is conserved, while the second one states that electric charge is conserved.